OCMu64: a solver for One-sided Crossing Minimization

### 2024-06-24 tex/description.tex
move 'no strong gluing' remark
add some refs

We write $u\b v$ when $u$ must come before $v$ in all minimal solutions, and say that $(u, v)$
is a \emph{fixed} pair. The following result is well known.

@@

\subparagraph{Gluing}
We now turn our attention to \emph{gluing}, i.e., proving that two vertices $u$
and $v$ always go right next to each other, and we can treat them as a single vertex. First,
let us see that we cannot get an analogous result to \cref{lem:sfimplf}.

\begin{remark}
  When $N(u)=N(v)$ in the unweighted case, or more generally $\bar u=\bar v$, we can glue $u$
    and $v$.
  Otherwise when $r(u,v)\leq 0$, there is an $X : A\to \mathbb R^{\geq 0}$ such that $(u, X, v)$ is
  strictly better than $(u,v,X)$ and $(X,u,v)$.
\end{remark}
This means that there is no `strong gluing'.
\begin{lemma}[Practical gluing]\label{pg}
    Let $u$ and $v$ satisfy $r(u, v) \leq 0$.
    A non-empty subset $X\subseteq B-\{u,v\}$ is \emph{blocking} when $c(u, X, v) \leq
    \min(c(u, v, X), c(X, u, v))$. If there is no blocking set, then we can glue $u, v$.
\end{lemma}
Again such sets $X$ can be found or proven to not exist using a knapsack
algorithm: add points $P_x = (r(u, x), r(x, v))$ and search for a non-empty
set summing to $\leq{}(0,0)$.

@@

\begin{description}
  \item[Graph simplification] We drop degree-$0$ vertices, merge identical
    vertices, and split the graph into \emph{independent} components when
    possible. We find an initial solution using local search that tries to move
    slices and optimally insert them, and re-label all nodes accordingly to make
    memory accesses more efficient.
  \item[Fixed pairs] We find all strongly fixed pairs and store them. For the
    exact track we also find practically fixed pairs. Instances for the parameterized
    track are simple enough that the overhead was not worth it. Also for each
    tail we search for new `tail-local' practically fixed pairs. In each state, we only
    try vertices $u\in T$ not fixed by another $v\in T$.
  \item[Gluing] We use the greedy strategy of \cref{greedy}. Our implementation
    of \cref{pg} contained a bug, so we did not use this. (Also benefits
    seemed limited.)
  \item[Tail cache] In each step, we search for the longest suffix of $T$ that
    was seen before, and reuse (the lower bound on) its excess. We also
    cache the tail-local practically fixed pairs.
  \item[Optimal insert] Instead of simply appending $u$ to $P$, we
    insert it in the optimal position. Note that the
    implementation is tricky because it interacts in complicated ways with
    the caching of results for each tail.
\end{description}
\end{document}

>>>>

We write $u\b v$ when $u$ must come before $v$ in all minimal solutions, and say that $(u, v)$
is a \emph{fixed} pair. The following result is well known \cite[Fact 4]{dujmovic_2004}.

@@

\subparagraph{Gluing}
We now turn our attention to \emph{gluing}, i.e., proving that two vertices $u$
and $v$ always go right next to each other, and we can treat them as a single vertex. First,
let us see that we cannot get `strong gluing' result analogous to \cref{lem:sfimplf}.

\begin{remark}
  When $N(u)=N(v)$ in the unweighted case, or more generally $\bar u=\bar v$, we can glue $u$
    and $v$.
  Otherwise when $r(u,v)\leq 0$, there is an $X : A\to \mathbb R^{\geq 0}$ such that $(u, X, v)$ is
  strictly better than $(u,v,X)$ and $(X,u,v)$.
\end{remark}
\begin{lemma}[Practical gluing]\label{pg}
    Let $u$ and $v$ satisfy $r(u, v) \leq 0$.
    A non-empty subset $X\subseteq B-\{u,v\}$ is \emph{blocking} when $c(u, X, v) \leq
    \min(c(u, v, X), c(X, u, v))$. If there is no blocking set, then we can glue $u, v$.
\end{lemma}
Again such sets $X$ can be found or proven to not exist using a knapsack
algorithm: add points $P_x = (r(u, x), r(x, v))$ and search for a non-empty
set summing to $\leq{}(0,0)$.

@@

\begin{description}
  \item[Graph simplification] We drop degree-$0$ vertices, merge identical
    vertices, and split the graph into \emph{independent} components when
    possible. We find an initial solution using the median heuristic \cite{eades_median,eades_median_2} and a local search that tries to move
    slices and optimally insert them \cite{makinen_experiments,eades_heuristcis}, and re-label all nodes accordingly to make
    memory accesses more efficient.
  \item[Fixed pairs] We find all strongly fixed pairs and store them. For the
    exact track we also find practically fixed pairs. Instances for the parameterized
    track are simple enough that the overhead was not worth it. Also for each
    tail we search for new `tail-local' practically fixed pairs. In each state, we only
    try vertices $u\in T$ not fixed by another $v\in T$.
  \item[Gluing] We use the greedy strategy of \cref{greedy}. Our implementation
    of \cref{pg} contained a bug, so we did not use this. (Also benefits
    seemed limited.)
  \item[Tail cache] In each step, we search for the longest suffix of $T$ that
    was seen before, and reuse (the lower bound on) its excess. We also
    cache the tail-local practically fixed pairs.
  \item[Optimal insert] Instead of simply appending $u$ to $P$, we
    insert it in the optimal position. Note that the
    implementation is tricky because it interacts in complicated ways with
    the caching of results for each tail.
\end{description}

\bibliography{bibliography}

\end{document}
